---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```
##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```


## What is mean total number of steps taken per day?
```{r}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```
##### 1. Make a histogram of the total number of steps taken each day
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency width 1000', binwidth=1000)
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
meanSteps <- mean(stepsByDay)
medianSteps <- median(stepsByDay)
```
* Mean: `r meanSteps`
* Median:  `r medianSteps`

## What is the average daily activity pattern?
```{r}
averageStepsInterval <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```
##### 1. Make a time series plot
```{r}
ggplot(data=averageStepsInterval, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute intervals") +
    ylab("avg steps") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps <- which.max(averageStepsInterval$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsInterval[mostSteps,'interval'])
```

* Most Steps at: `r timeMostSteps`


## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
missingValues <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r missingValues`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(Hmisc)
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r}
imputedStepsByDay <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(imputedStepsByDay, xlab='Total steps per day (Imputed)', ylab='Frequency using width 1000', binwidth=1000)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
imputedStepsByDayMean <- mean(imputedStepsByDay)
imputedStepsByDayMedian <- median(imputedStepsByDay)
```
* Mean (Imputed): `r imputedStepsByDayMean`
* Median (Imputed):  `r imputedStepsByDayMedian`




## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r}
imputedAveragedActivityData <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(imputedAveragedActivityData, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avg steps")
```
